Rotation Period Evolution in Low-Mass Binary Stars: The Impact of Tidal Torques and Magnetic Braking

### 2019-04-22 Paper/sync.tex

@@

\begin{figure*}[ht]
	\includegraphics[width=\textwidth]{../Plots/stellarExample.pdf}
   \caption{Stellar $R$ (left), $r_g$ (middle), and P$_{rot}$ (right) evolution for 0.2 M$_{\odot}$ (M, green), 0.7 M$_{\odot}$ (K, orange), and 1 M$_{\odot}$ (G, blue) mass stars computed according to \stellar, our interpolation of the \citet{Baraffe2015} stellar evolution models ($\S$~\ref{sec:methods:stellar}) combined with the \xxx{\citet{Matt2015} magnetic braking model}. Each dot denotes the approximate time when each star reaches the main sequence.}%
    \label{fig:stellarExample}%
\end{figure*}

@@

In both the CPL and CTL model, We assume $k_2 = 0.5$. This choice of $k_2$ does not impact our results as $k_2$ is degenerate with Q in the CPL model, e.g. the $k_2/Q$ scaling in Eq.~(\ref{eqn:cpl:z}), and with $\tau$ in the CTL model, e.g. $k_2 \tau$ scaling in Eq.~(\ref{eqn:ctl:z}), so we instead examine how our results scale with $Q$ and $\tau$.  Any constraints we derive as a function $Q$ or $\tau$ can trivially be scaled to other values of $k_2$.

@@

@@

\begin{figure*}
	\includegraphics[width=\textwidth]{../Plots/eqPerTwoPanelMatt.pdf}
   \caption{Evolution of stellar P$_{rot}$, normalized by P$_{eq}$ (see Eqn.~(\ref{eqn:cpl:eqPer}) and Eqn.~(\ref{eqn:ctl:eqPer}), for initial circular binary orbits according to the CPL (solid) and CTL (dashed) models with $Q = 10^6$ and $\tau = 0.1$ s, respectively, using the \citet{Matt2015} magnetic braking model. Left: P$_{rot}/$P$_{eq}$ for stars with P$_{orb}$ ranging from 5 d to 60 d. The black dotted line indicates the tidally-locked state where P$_{rot} = $ P$_{eq}$.  Right: Net P$_{rot}$ derivative due to stellar evolution, tidal torques, and magnetic braking.  We truncate each curve when the binary tidally locks. For P$_{orb} \leq 10$ d, both the CPL and CTL models predict that the binaries lock into synchronous rotation.  For all P$_{orb}$, the CPL models tidally-lock into synchronous rotation whereas the CTL model predicts subsynchronous (P$_{rot} > $ P$_{eq}$) rotation that persists for Gyrs.}%
    \label{fig:eqPer}%
\end{figure*}

@@

\begin{figure}
	\includegraphics[width=0.45\textwidth]{../Plots/eqPerShortPorbMatt.pdf}
   \caption{Evolution of stellar P$_{rot}$, normalized by P$_{eq}$ (see Eqn.~(\ref{eqn:cpl:eqPer}) and Eqn.~(\ref{eqn:ctl:eqPer}), for initial circular binary orbits with initial P$_{orb} = 5$ d according to the CPL (solid) and CTL (dashed) models for several values of $Q$ and $\tau$, respectively, using the \citet{Matt2015} magnetic braking model. Systems with strong tidal torques ($Q = 10^6$ or $\tau = 0.1$ s) tidally lock, whereas in systems with weaker tidal torques (larger $Q$ and smaller $\tau$, respectively), magnetic braking initially overpowers tidal torques, spinning down the stars past the tidally-locked state, resulting in subsynchronous rotation. }%
    \label{fig:eqPerShortPorb}%
\end{figure}

@@

\begin{figure*}[ht]
	\includegraphics[width=\textwidth]{../Plots/qTauPorbRatioHist.pdf}
   \caption{Median P$_{orb}/$P$_{rot}$ at the end of the simulation according to the CPL (left) and CTL (right) models binned by log$_{10}(Q)$ and log$_{10}(\tau)$, respectively, and P$_{orb}$.  For P$_{orb} < 10$ d, the CPL model predicts that most systems will tidally lock into synchronous rotation, regardless of $Q$, whereas the CTL model requires $\tau \geq 0.1$ s to tidally-lock.  Both models exhibit supersynchronous rotation (blue regions, P$_{rot} <$ P$_{orb}$) for P$_{orb} > 80$d, that do not in general reflect tidally-locking into a supersynchronous rotation, but rather at long P$_{orb}$, the combination of weak tidal torques and magnetic braking have not spun down the star enough for P$_{rot}$ to be close to P$_{orb}$ after 7 Gyr of evolution. For large $Q$ ($> 10^{6.5}$) and small $\tau$ ($\tau < 0.1$ s), weak tidal torques cannot prevent magnetic braking from spinning down stars past the tidally-locked state, producing a population of subsynchronous rotators (red regions, P$_{rot} >$ P$_{orb}$).}%
    \label{fig:qTauLock}%
\end{figure*}

@@

\begin{figure*}
	\includegraphics[width=\textwidth]{../Plots/porbProtPorbQHist.pdf}
   \caption{Median log$_{10}(Q)$ of primary stars binned by P$_{orb}$ and P$_{orb}/$P$_{rot}$ evolved using the CPL model.  In each bin, we compute the median, $16\%$, and $84\%$ percentiles for log$_{10}(Q)$ and display these values. Narrow constraints imply that certain tidal parameters are required to produce that rotation state, e.g. supersynchronous rotators near P$_{orb}/$P$_{rot} = 1.5$ require $Q \lsim 10^{5.5}$. }%
    \label{fig:qmap}%
\end{figure*}

@@

\begin{figure*}
	\includegraphics[width=\textwidth]{../Plots/porbProtPorbTauHist.pdf}
   \caption{Same format as Fig.~\ref{fig:qmap}, but for log$_{10}(\tau [s])$ under the CTL model. }%
    \label{fig:taumap}%
\end{figure*}

@@

\begin{figure}
	\includegraphics[width=0.5\textwidth]{../Plots/lockedCPL.pdf}
   \caption{Rotation state for tidally-locked (blue, P$_{rot} = $ P$_{eq}$), interacting (green, P$_{rot}$ within $10\%$ of P$_{eq}$ and not locked), and not locked (orange, remainder of binaries) stellar binaries. Left: P$_{rot}$ as a function of stellar mass and age according to our CPL simulations integrated to system ages uniformly sampled over $1-7$ Gyr. Right: Marginalized P$_{rot}$ distribution for each case.}%
    \label{fig:lockedCPL}%
\end{figure}

Both models predict a substantial population of tidally-locked fast rotators with P$_{rot} \lsim 20$ d, with tidally-locked stars systematically rotating faster (median CPL, CTL P$_{rot} = 27.6$ d and $7.4$ d) than unlocked (median CPL, CTL P$_{rot} = 41.3$ d and $40.8$ d) binaries. More massive stars are more likely to tidally-lock compared to less massive stars as tidal torques scale as $R^5$, and $R$ increases with stellar mass.  This feature is seen in the enhanced density of locked systems at larger masses for both tidal models, but in particular for the CPL model.

@@

\begin{figure}
	\includegraphics[width=0.5\textwidth]{../Plots/lockedCTL.pdf}
   \caption{Same format as Fig.~\ref{fig:lockedCPL}, but for the CTL simulations. }%
    \label{fig:lockedCTL}%
\end{figure}

@@

@@

\begin{figure*}[t]
	\includegraphics[width=\textwidth]{../Plots/protDist.pdf}
   \caption{P$_{rot}$ as a function of stellar mass and age according to our CPL (left), CTL (left center), and single star (right center) simulations integrated to system ages uniformly sampled over $1-7$ Gyr using the \citet{Matt2015} magnetic braking model. For each case, we only plot 2,500 systems for clarity but account for all systems when computing the marginalized distributions. Right: The P$_{rot}$ distribution for each case, marginalized over stellar mass.}%
    \label{fig:protDist}%
\end{figure*}

There is a clear age-dependence of P$_{rot}$ seen in the single star population with older stars rotating more slowly, a trend that is borne out in nature and is the critical assumption of gyrochronology methods that link P$_{rot}$ to stellar ages via the magnetic braking-driven long-term spin down of low-mass stars \citep[e.g.][]{Skumanich1972,Barnes2003,Barnes2007,Mamajek2008,Barnes2010,Meibom2015}. The only rapidly rotating single stars are young (ages $< 2$ Gyr) M dwarfs (M$ < 0.5$ M$_{\odot}$) that have just reached the main sequence and have not yet experienced substantial angular momentum loss due to magnetic braking.  In stark contrast, both tidal models predict a substantial population of tidally-locked fast rotators (P$_{rot} < 20$ d). Under the CPL and CTL models, $18\%$ and $19\%$ of simulated stars have P$_{rot} < 20 $ d, respectively, compared to only $5\%$ in the single star models.  At the long P$_{rot}$ end, the CPL model predicts a population of slow rotators with P$_{rot} \gsim 70$ d, a feature not seen in either the CTL or single star models, and is produced by tidal-locking in long P$_{orb}$ binaries with low tidal $Qs$, e.g. Fig.~\ref{fig:lockedCPL}.

@@

Our models predict that age does not strongly correlate with P$_{rot}$ in low-mass binaries, especially for P$_{orb} < 20 $ d. \xxx{The median ages and $68\%$ interval for fast rotators (P$_{rot} \lsim 20$ d) are $2.8^{+2.9}_{-1.5}$ Gyr and $3.1^{+2.6}_{-1.8}$ Gyr according to the CPL and CTL models, respectively}, compared to much younger single stars with ages of $1.3^{+0.5}_{-0.3}$ Gyr. We highlight this stark dichotomy between the age distribution of single and double stars in Fig.~\ref{fig:protAgeHist} by plotting a histogram of system ages from Fig.~\ref{fig:protDist} for single or primary stars in binaries with P$_{rot} < 20$ d.

\begin{figure}[h]
	\includegraphics[width=0.45\textwidth]{../Plots/protAgeHist.pdf}
   \caption{Histogram of rapidly-rotating (P$_{rot} < 20$ d) stellar ages for single and primary stars in binaries from Fig.~\ref{fig:protDist}. Rapidly-rotating single stars must be young (ages $\lsim 2$ Gyr), while tidally-locked rapidly-rotating binaries exhibit a wide range of ages.}%
    \label{fig:protAgeHist}%
\end{figure}

Tidal torques pose a fundamental problem for inferring ages of stars via \xxx{gyrochronology}. Regardless of the choice of equilibrium tidal model or magnetic braking model, stellar binaries readily tidally-lock, or at least strongly tidally-interact, across a wide range of P$_{orb}$ and primary star masses, decoupling P$_{rot}$ from age. For example, if one observed a rapidly rotating star with P$_{rot} \lsim 20$ d, gyrochronology models would predict ages $\lsim 2$ Gyr. If a rapidly-rotating star is actually an unresolved binary, it would likely be tidally-locked, decoupling P$_{rot}$ from age, causing the predictions of gyrochronology models to fail as its age could be at least several Gyrs, e.g. Fig.~\ref{fig:protAgeHist}. This effect is most likely to manifest in rapid rotators (P$_{rot} < 20$ d), but persists across all P$_{rot}$ up to 100 d, producing a significant contaminating signal, \textbf{e.g. Fig.~\ref{fig:protDist}}.

In general, it is difficult to accurately determine if a source is single star or a stellar binary via longterm photometric monitoring as only a small fraction of stars in binaries will occult one another. Observations of the binarity of field stars by \citet{Raghavan2010} and \citet{Duchene2013} indicate that roughly half of stars are in stellar binaries, with $10\%$ of these binaries having P$_{orb} \lsim 100$ d, suggesting that unless one accounts for binarity, stellar binaries will produce a non-negligible contaminating signal in any study of stellar rotation periods and any ages inferred via \xxx{gyrochronology} are potentially subject to systematic errors.  Moreover, this problem could be more significant as \citet{Simonian2018} found that most rapid rotators with P$_{rot} \leq 7.5$ d in the \kepler field are likely photometric binaries, suggesting that binary contamination in P$_{rot}$ studies could be widespread. We caution that any application of \xxx{gyrochronology} methods to predict ages for stars, especially those with P$_{rot} \lsim 20$ d, should rule out or account for stellar binarity, or otherwise risk deriving systematically incorrect ages.

@@

We compare our simulation results using both tidal models and the \citet{Matt2015} magnetic braking model to P$_{rot}$ measurements of primary stars in \kepler low-mass eclipsing stellar binaries by \citet{Lurie2017}. The purpose of this comparison is to see if our model predictions, which by design populate a wide, but physically-plausible, region of parameter space, can reproduce features observed in the data.  \citet{Lurie2017} measured 816 rotation periods for primary stars in \kepler EBs with star spot modulations and visually inspected each light curve to ensure their accuracy. The \citet{Lurie2017} dataset is the largest homogenous set of P$_{rot}$ measurements available for low-mass stellar binaries and represents the state of the art benchmark for studies of the influence of tides on P$_{rot}$ in stellar binaries. We compare our results to the P$_{1,min}$ P$_{rot}$ values reported by \citet{Lurie2017} as the authors demonstrated that these values are likely to be close to the equatorial P$_{rot}$ that we track in our simulations. In Fig.~\ref{fig:lurie7}, we display P$_{orb}/$P$_{rot}$ as a function of P$_{orb}$ for both the CPL and CTL models where each simulation was integrated to an age uniformly sampled over $1-7$ Gyr, consistent with ages of \kepler field stars \citep{Chaplin2014}, and the \citet{Lurie2017} data.

>>>>

\subsubsection{\xxx{Core-Envelope Coupling}} \label{sec:methods:twoLayer}

\xxx{Our simplified model assumes that stars follow solid body rotation, whereas in low-mass stars, coupling between the radiative core and convective envelope can impact the surface rotation period evolution modeled in this work \citep{MacGregor1991,Allain1998}. Recent work by \citet{Gallet2013} and \citet{Gallet2015} find that the rapidly-rotating stellar core acts as an angular momentum reservoir for the convective envelope, potentially transferring angular momentum within the stellar interior, i.e. into the envelope, for up to 1 Gyr, although this process depends strongly on the adopted magnetic braking model, the initial rotation rate, and the stellar mass \citep{Irwin2007,Bouvier2008,Irwin2009}. This core-to-envelope angular momentum transfer can counteract the spin-down caused by magnetic braking on the main sequence. For a given magnetic braking model, this angular momentum transfer within the stellar interior would likely lead to slightly faster stellar rotation on the early main sequence and a wider spread of rotation periods on the zero-age main sequence than is predicted by solid body rotation models \citep{Gallet2013}. This effect could be more pronounced for low-mass M-dwarfs, potentially accounting for their long spin-down timescales \citep{Gallet2015}. Modeling core-envelope coupling is beyond the scope of this work, however.}

\subsubsection{\xxx{Example Stellar Evolution}} \label{sec:methods:exampleStellar}

@@

\begin{figure*}[ht]
	\includegraphics[width=\textwidth]{../Plots/stellarExample.pdf}
   \caption{\xxx{Updated Figure} Stellar $R$ (left), $r_g$ (middle), and P$_{rot}$ (right) evolution for 0.2 M$_{\odot}$ (M, green), 0.7 M$_{\odot}$ (K, orange), and 1 M$_{\odot}$ (G, blue) mass stars computed according to \stellar, our interpolation of the \citet{Baraffe2015} stellar evolution models ($\S$~\ref{sec:methods:stellar}) combined with the \xxx{\citet{Matt2015} magnetic braking model}. Each dot denotes the approximate time when each star reaches the main sequence.}%
    \label{fig:stellarExample}%
\end{figure*}

@@

In both the CPL and CTL model, We assume $k_2 = 0.5$. This choice of $k_2$ does not impact our results as $k_2$ is degenerate with Q in the CPL model, e.g. the $k_2/Q$ scaling in Eq.~(\ref{eqn:cpl:z}), and with $\tau$ in the CTL model, e.g. $k_2 \tau$ scaling in Eq.~(\ref{eqn:ctl:z}), so we instead examine how our results scale with $Q$ and $\tau$.  Any constraints we derive as a function $Q$ or $\tau$ can trivially be scaled to other values of $k_2$. \xxx{For example, a common re-parameterization of Q is the reduced tidal quality factor, $Q' = 3Q/2k_2$ \citep[e.g.][]{Leconte2010}. Given our choice of $k_2 = 0.5$, this reduces to $Q' = 3 Q$.}

@@

\subsubsection{\xxx{The Dynamical Tide}} \label{sec:methods:dynamicalTide}

\xxx{An additional mechanism for tidal dissipation in low-mass stellar binaries not considered in this work is the dynamical tide. Dynamical tidal dissipation arises from the turbulent viscous damping of inertial waves that are excited in the stellar convective envelope by a tidal perturber, with Coriolis acceleration serving as the restoring force \citep{Zahn1975, Ogilvie2007}. Under the dynamical tide formalism, the stellar mass, evolving stellar structure, rotation rate, and tidal forcing frequency can all strongly impact the strength of tidal dissipation, which can span many orders of magnitude \citep{Ogilvie2007,Ogilvie2013,Mathis2015,Gallet2017}. For example, adopting the tidal frequency-averaged model for tidal dissipation of \citet{Ogilvie2013}, both \citet{Mathis2015} and \citet{Gallet2017} show that dynamical tidal dissipation is enhanced on the pre-main sequence via the expansion of the stellar radiative core and rapid rotation, whereas the magnetic braking-driven spin-down on the main sequence decreases the tidal dissipation. Dissipation due to the dynamical tide could be important for some of the systems considered in this work since, for coplanar binary stars on circular orbits, inertial waves are excited in the stellar convective envelopes for P$_{orb} > $P$_{rot}/2$ and can drive significant spin and orbital evolution \citep[e.g.][]{Witte2002,Ogilvie2007,Bolmont2016}. Although semi-analytic models for dynamical tidal dissipation that account for the evolving stellar structure and rotation exist \citep[e.g.][]{Ogilvie2013,Mathis2015,Bolmont2016,Gallet2017}, we do not consider them here as they are currently limited to circular orbits. We instead focus on exploring the impact of the equilibrium tide across a wide range of parameter space and leave an examination of how the combination of the dynamical and equilibrium tide impacts the rotation period evolution of low-mass binary stars for future work.}

@@

\begin{figure*}
	\includegraphics[width=\textwidth]{../Plots/eqPerTwoPanelMatt.pdf}
   \caption{\xxx{Updated Figure} Evolution of stellar P$_{rot}$, normalized by P$_{eq}$ (see Eqn.~(\ref{eqn:cpl:eqPer}) and Eqn.~(\ref{eqn:ctl:eqPer}), for initial circular binary orbits according to the CPL (solid) and CTL (dashed) models with $Q = 10^6$ and $\tau = 0.1$ s, respectively, using the \citet{Matt2015} magnetic braking model. Left: P$_{rot}/$P$_{eq}$ for stars with P$_{orb}$ ranging from 5 d to 60 d. The black dotted line indicates the tidally-locked state where P$_{rot} = $ P$_{eq}$.  Right: Net P$_{rot}$ derivative due to stellar evolution, tidal torques, and magnetic braking.  We truncate each curve when the binary tidally locks. For P$_{orb} \leq 10$ d, both the CPL and CTL models predict that the binaries lock into synchronous rotation.  For all P$_{orb}$, the CPL models tidally-lock into synchronous rotation whereas the CTL model predicts subsynchronous (P$_{rot} > $ P$_{eq}$) rotation that persists for Gyrs.}%
    \label{fig:eqPer}%
\end{figure*}

@@

\begin{figure}
	\includegraphics[width=0.45\textwidth]{../Plots/eqPerShortPorbMatt.pdf}
   \caption{\xxx{Updated Figure} Evolution of stellar P$_{rot}$, normalized by P$_{eq}$ (see Eqn.~(\ref{eqn:cpl:eqPer}) and Eqn.~(\ref{eqn:ctl:eqPer}), for initial circular binary orbits with initial P$_{orb} = 5$ d according to the CPL (solid) and CTL (dashed) models for several values of $Q$ and $\tau$, respectively, using the \citet{Matt2015} magnetic braking model. Systems with strong tidal torques ($Q = 10^6$ or $\tau = 0.1$ s) tidally lock, whereas in systems with weaker tidal torques (larger $Q$ and smaller $\tau$, respectively), magnetic braking initially overpowers tidal torques, spinning down the stars past the tidally-locked state, resulting in subsynchronous rotation. }%
    \label{fig:eqPerShortPorb}%
\end{figure}

@@

\begin{figure*}[ht]
	\includegraphics[width=\textwidth]{../Plots/qTauPorbRatioHist.pdf}
   \caption{\xxx{Updated Figure} Median P$_{orb}/$P$_{rot}$ at the end of the simulation according to the CPL (left) and CTL (right) models binned by log$_{10}(Q)$ and log$_{10}(\tau)$, respectively, and P$_{orb}$.  For P$_{orb} < 10$ d, the CPL model predicts that most systems will tidally lock into synchronous rotation, regardless of $Q$, whereas the CTL model requires $\tau \geq 0.1$ s to tidally-lock.  Both models exhibit supersynchronous rotation (blue regions, P$_{rot} <$ P$_{orb}$) for P$_{orb} > 80$d, that do not in general reflect tidally-locking into a supersynchronous rotation, but rather at long P$_{orb}$, the combination of weak tidal torques and magnetic braking have not spun down the star enough for P$_{rot}$ to be close to P$_{orb}$ after 7 Gyr of evolution. For large $Q$ ($> 10^{6.5}$) and small $\tau$ ($\tau < 0.1$ s), weak tidal torques cannot prevent magnetic braking from spinning down stars past the tidally-locked state, producing a population of subsynchronous rotators (red regions, P$_{rot} >$ P$_{orb}$).}%
    \label{fig:qTauLock}%
\end{figure*}

@@

\begin{figure*}
	\includegraphics[width=\textwidth]{../Plots/porbProtPorbQHist.pdf}
   \caption{\xxx{Updated Figure} Median log$_{10}(Q)$ of primary stars binned by P$_{orb}$ and P$_{orb}/$P$_{rot}$ evolved using the CPL model.  In each bin, we compute the median, $16\%$, and $84\%$ percentiles for log$_{10}(Q)$ and display these values. Narrow constraints imply that certain tidal parameters are required to produce that rotation state, e.g. supersynchronous rotators near P$_{orb}/$P$_{rot} = 1.5$ require $Q \lsim 10^{5.5}$. }%
    \label{fig:qmap}%
\end{figure*}

@@

\begin{figure*}
	\includegraphics[width=\textwidth]{../Plots/porbProtPorbTauHist.pdf}
   \caption{\xxx{Updated Figure} Same format as Fig.~\ref{fig:qmap}, but for log$_{10}(\tau [s])$ under the CTL model. }%
    \label{fig:taumap}%
\end{figure*}

@@

\begin{figure}
	\includegraphics[width=0.5\textwidth]{../Plots/lockedCPL.pdf}
   \caption{\xxx{Updated Figure} Rotation state for tidally-locked (blue, P$_{rot} = $ P$_{eq}$), interacting (green, P$_{rot}$ within $10\%$ of P$_{eq}$ and not locked), and not locked (orange, remainder of binaries) stellar binaries. Left: P$_{rot}$ as a function of stellar mass and age according to our CPL simulations integrated to system ages uniformly sampled over $1-7$ Gyr. Right: Marginalized P$_{rot}$ distribution for each case.}%
    \label{fig:lockedCPL}%
\end{figure}

Both models predict a substantial population of tidally-locked fast rotators with P$_{rot} \lsim 20$ d, with tidally-locked stars systematically rotating faster (median CPL, CTL P$_{rot} = 27.6$ d and $7.4$ d) than unlocked (median CPL, CTL P$_{rot} = 41.3$ d and $40.8$ d) binaries. More massive stars are more likely to tidally-lock compared to less massive stars as tidal torques scale \xxx{with the stellar masses and as $R^5$, with $R$ increasing with stellar mass.  This feature is seen in the enhanced density of locked systems at larger masses for both tidal models, but in particular for the CPL model. We highlight this enhanced density of locked binaries in systems with more massive primaries in the marginalized mass distributions in the top panels of Fig.~\ref{fig:lockedCPL} and Fig.~\ref{fig:lockedCTL}.}

@@

\begin{figure}
	\includegraphics[width=0.5\textwidth]{../Plots/lockedCTL.pdf}
   \caption{\xxx{Updated Figure} Same format as Fig.~\ref{fig:lockedCPL}, but for the CTL simulations. }%
    \label{fig:lockedCTL}%
\end{figure}

@@

\begin{figure*}[t]
	\includegraphics[width=\textwidth]{../Plots/protDist.pdf}
   \caption{\xxx{Updated Figure} P$_{rot}$ as a function of stellar mass and age according to our CPL (left), CTL (left center), and single star (right center) simulations integrated to system ages uniformly sampled over $1-7$ Gyr using the \citet{Matt2015} magnetic braking model. For each case, we only plot 2,500 systems for clarity but account for all systems when computing the marginalized distributions. Right: The P$_{rot}$ distribution for each case, marginalized over stellar mass.}%
    \label{fig:protDist}%
\end{figure*}

\xxx{There is a clear age-dependence of P$_{rot}$ seen in the single star population with older stars rotating more slowly, a trend that is borne out in nature and is the critical assumption of gyrochronology methods that link P$_{rot}$ to stellar ages via the magnetic braking-driven long-term spin down of low-mass stars \citep[e.g.][]{Skumanich1972,Barnes2003,Barnes2007,Mamajek2008,Barnes2010,Meibom2015}.
In stark contrast, both tidal models predict that age does not always strongly correlate with P$_{rot}$ as binaries at a given P$_{rot}$ can assume a range of ages, especially at short and long P$_{rot}$, due to the influence of tidal torques. In binaries, tidal torques tend to drive the P$_{rot}$ evolution away from that of single stars and towards P$_{eq}$, either maintaining rapid rotation in tidally-locked short P$_{orb}$ systems, or working with magnetic braking to slow P$_{rot}$ beyond that of single stars of the same age.}

\xxx{We quantify the impact of binarity on gyrochronology in Fig.~\ref{fig:gyro} by computing the percent difference between the mean ages of single and binary stars, for both tidal models, in mass and P$_{rot}$ bins for the populations depicted in Fig.~\ref{fig:protDist}. This quantity, referred to here as the ``Relative Age Error", represents the systematic error incurred by assigning a binary star the age expected for single stars at a given mass and P$_{rot}$. For this comparison, we select the subset of tidally-locked and tidally-interacting binaries as classified in Fig.~\ref{fig:lockedCPL} and Fig.~\ref{fig:lockedCTL}, accounting for $31\%$ and $24\%$ of CPL and CTL binaries, respectively, as the remaining binaries weakly tidally-interact, and hence their P$_{rot}$ evolution will predominately follow that of single stars.}

\begin{figure*}[t]
	\includegraphics[width=\textwidth]{../Plots/gyro.pdf}
   \caption{\xxx{New Figure: Relative age error between single and binary stars for both the CPL (left) and CTL (right) tidal models. The error is computed as the percent difference between the mean ages of single and binary stars in mass and P$_{rot}$ bins.}}%
    \label{fig:gyro}%
\end{figure*}

\xxx{Gyrochronology ages, when applied to tidally-interacting binaries systematically underestimate the ages, with the age error increasing with decreasing P$_{rot}$ and increasing with primary star mass where tides are stronger. For binaries with P$_{rot} \lsim 30 d$, gyrochronology ages are underestimated by $50\%$, with this error growing to $300\%$ for P$_{rot} \approx 10$ d. For slow rotators with P$_{rot} \gsim 40$ d, gyrochronology ages are slightly overestimated by up to $25\%$ for near solar-mass stars. Single star-only models fail to produce near-solar mass stars with P$_{rot} \lsim 10$ d nor P$_{rot} \gsim 50$ as only tidally-interacting binaries at short and long P$_{orb}$, respectively, can produce such rotation states, so we can not compare binary star and single star ages there.}

\xxx{... This effect is most pronounced at short P$_{orb}$, and hence P$_{rot}$, where tidal torques in binaries are strongest, leading to relative age errors of up to $300\%$ for both tidal models relative to single stars. This effect would be more pronounced for the most rapidly-rotating stars, e.g. tidally-interacting near-solar mass binaries with P$_{rot} \lsim 10$ d, however our single star model and age prior cannot produce such rapid rotators as only those with ages $\lsim 1$ Gyr could have P$_{rot} \lsim 10$ d. In Fig.~\ref{fig:protAgeHist}, we further examine the age distributions of rapid rotators.}

\xxx{The age distribution of tidally-interacting binaries with P$_{rot} < 20 $ d is inconsistent with that of single stars. In this range, the median ages and $68\%$ interval are $2.3^{+2.9}_{-0.9}$ Gyr and $2.4^{+3.0}_{-1.1}$ Gyr according to the CPL and CTL models, respectively, compared to the much younger single stars with ages of $1.6^{+0.8}_{-0.4}$ Gyr. We highlight this dichotomy in Fig.~\ref{fig:protAgeHist} by plotting a histogram of system ages from Fig.~\ref{fig:protDist} for single or primary stars in binaries with P$_{rot} < 20$ d.}

\begin{figure}[h]
	\includegraphics[width=0.45\textwidth]{../Plots/protAgeHist.pdf}
   \caption{\xxx{Updated Figure} Histogram of rapidly-rotating (P$_{rot} < 20$ d) stellar ages for single and primary stars in binaries from Fig.~\ref{fig:protDist}. Rapidly-rotating single stars must be young (ages $\lsim 2$ Gyr), while tidally-locked rapidly-rotating binaries exhibit a wide range of ages.}%
    \label{fig:protAgeHist}%
\end{figure}

Tidal torques pose a fundamental problem for inferring ages of stars via \xxx{gyrochronology}. Regardless of the choice of equilibrium tidal model or magnetic braking model, stellar binaries readily tidally-lock, or at least strongly tidally-interact, across a wide range of P$_{orb}$ and primary star masses, decoupling P$_{rot}$ from age. For example, if one observed a rapidly rotating star with P$_{rot} \lsim 20$ d, gyrochronology models would predict ages \xxx{$\lsim 1.6$} Gyr. If a rapidly-rotating star is actually an unresolved binary, it would likely be tidally-locked, decoupling P$_{rot}$ from age, causing the predictions of gyrochronology models to fail as its age could be at least several Gyrs, e.g. Fig.~\ref{fig:protAgeHist}. This effect is most likely to manifest in rapid rotators (P$_{rot} < 20$ d), but persists across all P$_{rot}$ up to 100 d, producing a significant contaminating signal, \xxx{e.g. Fig.~\ref{fig:protDist} and Fig.~\ref{fig:gyro}}.

In general, it is difficult to accurately determine if a source is single star or a stellar binary via longterm photometric monitoring as only a small fraction of stars in binaries will occult one another. Observations of the binarity of field stars by \citet{Raghavan2010} and \citet{Duchene2013} indicate that roughly half of stars are in stellar binaries, with $10\%$ of these binaries having P$_{orb} \lsim 100$ d, suggesting that unless one accounts for binarity, stellar binaries will produce a non-negligible contaminating signal in any study of stellar rotation periods and any ages inferred via \xxx{gyrochronology} are potentially subject to systematic errors.  Moreover, this problem could be more significant as \citet{Simonian2018} found that most rapid rotators with P$_{rot} \leq 7.5$ d in the \kepler field \xxx{are consistent with tidally-synchronized photometric binaries, suggesting that binary contamination in P$_{rot}$ studies could be widespread}. We caution that any application of \xxx{gyrochronology} methods to predict ages for stars, especially those with P$_{rot} \lsim 20$ d, should rule out or account for stellar binarity, or otherwise risk deriving systematically incorrect ages.

@@

% cut
% There is a clear age-dependence of P$_{rot}$ seen in the single star population with older stars rotating more slowly, a trend that is borne out in nature and is the critical assumption of gyrochronology methods that link P$_{rot}$ to stellar ages via the magnetic braking-driven long-term spin down of low-mass stars \citep[e.g.][]{Skumanich1972,Barnes2003,Barnes2007,Mamajek2008,Barnes2010,Meibom2015}. The only rapidly rotating single stars are young (ages $< 2$ Gyr) M dwarfs (M$ < 0.5$ M$_{\odot}$) that have just reached the main sequence and have not yet experienced substantial angular momentum loss due to magnetic braking.  In stark contrast, both tidal models predict a substantial population of tidally-locked fast rotators (P$_{rot} < 20$ d). Under the CPL and CTL models, $18\%$ and $19\%$ of simulated stars have P$_{rot} < 20 $ d, respectively, compared to only $5\%$ in the single star models.  At the long P$_{rot}$ end, the CPL model predicts a population of slow rotators with P$_{rot} \gsim 70$ d, a feature not seen in either the CTL or single star models, and is produced by tidal-locking in long P$_{orb}$ binaries with low tidal $Qs$, e.g. Fig.~\ref{fig:lockedCPL}.

@@

@@

We compare our simulation results using both tidal models and the \citet{Matt2015} magnetic braking model to P$_{rot}$ measurements of primary stars in \kepler low-mass eclipsing stellar binaries by \citet{Lurie2017}. The purpose of this comparison is to see if our model predictions, which by design populate a wide, but physically-plausible, region of parameter space, can reproduce features observed in the data.  \citet{Lurie2017} measured 816 rotation periods for primary stars in \kepler EBs with star spot modulations and visually inspected each light curve to ensure their accuracy. The \citet{Lurie2017} dataset is the largest homogenous set of P$_{rot}$ measurements available for low-mass stellar binaries and represents the state of the art benchmark for studies of the influence of tides on P$_{rot}$ in stellar binaries. We compare our results to the P$_{1,min}$ P$_{rot}$ values reported by \citet{Lurie2017} as the authors demonstrated that these values are likely to be close to the equatorial P$_{rot}$ that we track in our simulations. In Fig.~\ref{fig:lurie7}, we display P$_{orb}/$P$_{rot}$ as a function of P$_{orb}$ for both the CPL and CTL models where each simulation was integrated to an age uniformly sampled over $1-7$ Gyr, consistent with ages of \kepler field stars \citep{Chaplin2014}.

Both tidal models qualitatively reproduce most features seen in the \citet{Lurie2017} data, especially at short P$_{orb}$. Nearly all simulated binaries with P$_{orb} < 4$ d have circularized orbits and synchronized spins due to strong tidal torques at short stellar separations, in agreement with the \citet{Lurie2017} observations. At very short P$_{orb}$, in the absence of a perturbing tertiary companion, circularization and synchronization is the inevitable end state for low-mass binaries, in good agreement with theoretical predictions \citep{Counselman1973}. For P$_{orb} \gsim 4$ d, our models produce large populations of subsynchronous rotators. Although \citet{Lurie2017} argues that differential rotation creates the subsynchronous population, we find that the competition between weak tidal torques and magnetic braking described in $\S$~\ref{sec:eq} naturally produces subsynchronous rotators.  For example under the CTL model, our simulated subsynchronous binaries have systematically weaker tidal torques than the rest of the sample, as expected, as they have a median $\tau = 0.17$ s, compared to the median $\tau = 0.47$ s for the tidally-locked supersynchronous population.  
\begin{figure*}[t]
	\includegraphics[width=\textwidth]{../Plots/lurieFig7.pdf}
   \caption{\xxx{Updated Figure} P$_{orb}/$P$_{rot}$ as a function of P$_{orb}$ according to the CPL model (left) and the CTL model (middle), and \citet{Lurie2017} \kepler EB observations (right). All points are colored by $e$.  In the right panel, the \kepler EBs at low P$_{orb}$ and low P$_{orb}/$P$_{rot}$ are likely either brown dwarfs or exoplanets \citep{Lurie2017}, and hence are not modeled by our simulations, so we do not consider them, but we display them for completeness.}%
    \label{fig:lurie7}%
\end{figure*}
